---
title: "Initial Korean covid 19"
author: Sarah Castellon
date: October 10 2024
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
kRdata <- kRdata %>%
mutate(across(where(is.character), as.factor))

```
### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 2020
### This data was retrieved from a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   

# Methods  

$H_0$ There is no difference in dying depending on sex  
$H_A$ There is a difference in dying depending on sex

The data is from Korea in 2022  

result~sex
this is categorical ~ categorical 
We will use the helper file cat~cat 
  
# is dying of covid independent of sex?



## Graphical results - use ggplot commands to do this  

```{r}
barchartGC(~result + Sex,data=kRdata)
barchartGC(~result + Sex,data=kRdata, type="percent")

```

The bar chart show that more males died than women. 


## Numerical results

```{r}
table1 <- xtabs(~result + Sex, data=kRdata)
rowPerc(table1)
colPerc(table1)
```

Among the people who died  there was 64% of men who died and 36% of women who died.  

## Inferential results 

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

the p value for the Chi squared test is p-value = 0.001 and the Fisher exact test has a p-value of .00009. The odd of the Fisher test also states that men are 2.3 more acceptable of dying from covid 19. Therefore since both the p-values are lower then 0.05 we reject the null hypothesis. 

## Conclusions  

Question: is dying of covid independent of sex?

$H_0$ There is no difference in dying depending on sex  
$H_A$ There is a difference in dying depending on sex

The data is from Korea in 2022  

result~sex
this is categorical ~ categorical 
We will use the helper file cat~cat 

The data is from Korea in 2022  

We reject the null hypothesis since, 
The bar chart show that more males died than women.   
The numerical results show the p value for the Chi squared test is p-value is  0.001 and the Fisher exact test has a p-value of .00009. The odd of the Fisher test also states that men are 2.3 more acceptable of dying from covid 19. Therefore since both the p-values are lower then 0.05 we reject the null hypothesis. 
